---
title: "Homework 3"
format:
  html: 
    toc: true # includes a table of contents in rendered html format
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

# Set up

## Reading in packages
```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

## Data cleanup
```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

# Model code
```{r}
# Null model
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame

# Model 1: total biomass as a function of SLA, water treatment, and species
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)
par(mfrow = c(2, 2))

# Model 2: total biomass as a function of SLA and water treatment
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

# Model 3: total biomass as a function of SLA and species
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

# Model 4: total biomass as a function of water treatment and species
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)
# Determines best model
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)
# plots model 4 to use in question 1b
plot(model4)

```

# Problem 1
## a. 
```{r}

# table
## puts models into a list
models <- list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4)

modelsummary(models, # this function takes a list of models
             title = "Table 1: Multiple linear regression of the 5 models. The rows showcase the predictors for each model. The columns show each numbered model." # table title
             ) 


```

## b.
**To examine the influence of specific leaf area, water treatment, and species on total biomass, I created 5 models. The null model was a function of none of the predictors on total biomass, while model 1 was a function of all of the predictors on total biomass. Models 2, 3, and 4 were functions of 2 different predictors on total biomass. To determine the model that best described the influences on total biomass, I used multiple linear regression and found that model 4 (total biomass as a function of water treatment and species) had the lowest AIC value (-159.2). To evaluate linear model assumptions, I plotted model 4 and found that it was linear, normally distributed, homoscedastic, and has no high leverage points.**

## c.



