---
title: "Homework 3"
format: 
  html: docx
    toc: true # includes a table of contents in rendered html format
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

# Set up

## Reading in packages
```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)
library(ggplot2)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

## Data cleanup
```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

# Model code
```{r}
# Null model
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame

# Model 1: total biomass as a function of SLA, water treatment, and species
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)
par(mfrow = c(2, 2))

# Model 2: total biomass as a function of SLA and water treatment
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

# Model 3: total biomass as a function of SLA and species
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

# Model 4: total biomass as a function of water treatment and species
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)
# Determines best model
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)
# plots model 4 to use in question 1b
plot(model4)

```

# Problem 1
## a. 
```{r}

# table
## puts models into a list
models <- list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4)

modelsummary(models, # this function takes a list of models
             title = "Table 1: Multiple linear regression of the 5 models. The rows showcase the predictors for each model. The columns show each numbered model." # table title
             ) 


```

## b.
**To examine the influence of specific leaf area, water treatment, and species on total biomass, I created 5 models. The null model was a function of none of the predictors on total biomass, while model 1 was a function of all of the predictors on total biomass. Models 2, 3, and 4 were functions of 2 different predictors on total biomass. To determine the model that best described the influences on total biomass, I used multiple linear regression and found that model 4 (total biomass as a function of water treatment and species) had the lowest AIC value (-159.2). To evaluate linear model assumptions, I plotted model 4 and found that it was linear, normally distributed, homoscedastic, and has no high leverage points.**

## c + d.
```{r}
ggpredict(model4, 
          terms = c("water_treatment", 
                    "species_name")) %>% 
  plot(show_data = TRUE,
       facets = TRUE,
       show_legend = FALSE,
      colors = "ipsum") +
  theme_classic() +
  labs(title = "Total biomass as a function of water treatment and species",
       x = "Water Treatment",
       y = "Total Biomass (g)",
      caption = str_wrap("Figure 1. Total biomass as a function of water treatment and species. The darkest dot represents the model prediction and the transparent dots represent the underlying data. Data source: Valliere JM, Zhang J, Sharifi MR, Rundel PW. Can we condition native plants to increase drought tolerance and improve restoration success? Ecol Appl. 2019 Apr;29(3):e01863. doi: 10.1002/eap.1863. Epub 2019 Mar 4. PMID: 30831005."))
```

## e.
**Water treatment and species were the predictors that best described total mass (AIC = -159.2). On average, well-watered plants had a higher total biomass than drought stressed plants. On average Grindelia camporum had the highest total biomass with Eschscholzia californica, Nasella pulchra, and Encelia californica slightly smaller than it. On average Acmispon glaber had the lowest total biomass wiht Penstemon centranthifolius and Salvia leucophylla slightly bigger than it.**

# Problem 2
## a.
**For my data, I could do a piece of art. I would first change my graph to be a line plot with number of times the water bottle was refilled on the y-axis and date on the x-axis. I am thinking that the line of data could be the water flowing into the bottle. I would then draw a bottle below it to catch the water. This would be affective visualization because it includes the data and art that resembles what the data is of.**

# Problem 3
## a.
**This study measured the probability of food insecurity among people with disabilities. There was a focus on number of disabilities and type of disabilities. The author uses a logistical regression to model the probability of any and severe food insecurity as a function of the number of disabilities and the type of disability, controlling for socio-demographic variables.**
**INSERT FIGURE 1**

## b. 
**The first graph is a line plot which shows the probability of food insecurity (y-axis) based on the number of disabilities (x-axis). The second graph is a  histogram plot which shows the probability of food insecurity (y-axis) based on the category of disability (x-axis). The author's choice to separate these graphs made sense since number of disabilities is continuous data and the category of disability is categorical data. Both graphs show the mean probability of food insecurity and the standard error, but do not show the underlying data.**

## c.
**The authors did a good job handling 'visual clutter.' They graphs are relatively simple and primarily focus on the probabilities for each x value. The data:ink ratio is relatively high. There data is portrayed in a simple matter and is not cluttered or overwhelmed by grid lines and extraneous elements.**

## d.
**One recommendation I have is to change the y-axis values. Instead of probability represented as a decimal (ex: 0.1), it should be represented as a percentage (ex: 10%). This would make it easier to see on the graphs and easier for people with little research experience to understand. Another recommendation I have is to make the histogram bars a lighter and more transparent color on the inside (fill) and keep a dark outline for the bars. This would make the graph to seem more simple and the coloration of the graph would be less aggressive. I think this could help reduce visual clutter.**

